---
title: "Homework1_RBridge_MSDS2020"
author: "Bonnie Cooper"
date: "12/22/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Bridge Week 1 Assignment

Please create the following exercises in .rmd format, publish to rpub and submit both the rmd file and the rpub link


### 1. Write a loop that calculates 12-factorial
```{r}
#initialize variables
numFactorial <- 12 # 12!
factorialResult <- 1 # initialize our output as 1 since 0! = 1

#our n! factorial for loop
for( fact in 1:numFactorial )
{
  nextFact = fact * factorialResult
  factorialResult = nextFact
}
#display our result
cat( '12! = ', factorialResult )
```
### 2. Show how to create a numeric vector that contains the sequence from 20 to 50 by 5.

```{r}
#using the seq() fxn to specify the range from 20 to 50 in steps of 5
numVec <- seq( from = 20, to = 50, by = 5 )
#display our result
cat( 'numVec: ', numVec ) 
```
### 3. Create the function “quadratic” that takes a trio of input numbers a, b, and c and solve the quadratic
equation. The function should print as output the two solutions.
```{r}
quadratic <- function( a, b, c )
{
  positiveR <- ((-b ) + sqrt( ( b^2 ) - 4*a*c ) )/( 2*a )
  negativeR <- ((-b ) - sqrt( ( b^2 ) - 4*a*c ) )/( 2*a )
  cat( 'The positive root = ', positiveR, "\n" )
  #print( positiveR )
  cat( 'The negative root = ', negativeR )
  #print( negativeR )
}
#test cases for quadratic: 
# moar @ https://www.mesacc.edu/~scotz47781/mat120/notes/quad_formula/quad_formula_practice.html
#quadratic( 1,6,14 )
quadratic( 8,14,-15 )

```

