---
title: "R Bridge Week 2 BMC"
author: "Bonnie Cooper"
date: "12/25/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Bridge Course Week 2 Assignment

One of the challenges in working with data is wrangling. In this assignment we will use R to perform several tasks.
Here is a list of data sets: <http://vincentarelbundock.github.io/Rdatasets/> 
We will be selecting a set from this csv index list



### Selecting a CSV, "Statistics of Deadly Quarrels"
```{r}
#Statistics of Deadly Quarrels, 'Quarrels' from the HistData package
library( HistData )
#nrow( Quarrels )
#ncol( Quarrels )
head( Quarrels )
```

### Task 1:
Use the summary function to gain an overview of the data set. Then display the mean and median for at least two attributes

```{r}
#Using the summary function to gain an overview of the data set
#colnames( Quarrels )
#dim( Quarrels )
summary( Quarrels )
```
```{r}
#displaying the mean and median of 2 attributes

deaths_descriptiveSTATS <- list( Mean = mean( Quarrels$deaths ),
                                 Median = median(Quarrels$deaths ) ) 
months_descriptiveSTATS <- list( Mean = mean( Quarrels$months ),
                                 Median = median( Quarrels$months ) )
example_descriptiveSTATS <- list( DeathSTATs = deaths_descriptiveSTATS,
                                  MonthSTATs = months_descriptiveSTATS )
example_descriptiveSTATS


```

### Task #2:
Create a new data frame with a subset of the columns and rows. Make sure to rename it.

```{r}
subQuarrels <- Quarrels
subQuarrels <- subQuarrels[, c( 'year', 'international', 'colonial', 'revolution', 'deaths', 'months' ) ]
head( subQuarrels )
```

### Task #3:
Create new column names for the new data frame

```{r}
names( subQuarrels ) <- c( 'Year', 'International', 'Colonial',
                           'Revolution', 'Deaths', 'Months')
head( subQuarrels )
```

### Task #4:
Use the summary function to create an overview of your new data frame. Print the mean and median for the same two attributes. Please compare.

```{r}
summary( subQuarrels )
#class( subQuarrels )
deaths_descriptiveSTATS2 <- list( Mean = mean( subQuarrels$Deaths ),
                                  Median = median(subQuarrels$Deaths)) 
months_descriptiveSTATS2 <-list(Mean=mean(subQuarrels$Months ),
                                 Median = median( subQuarrels$Months ))
example_descriptiveSTATS2 <-list( DeathSTATs =deaths_descriptiveSTATS2,
                                  MonthSTATs =months_descriptiveSTATS2)
#display the descriptive stats for the same 2 features (deaths & months)
example_descriptiveSTATS2

#use the function 'identical' to directly compare the same descriptive stats from the previous calculation.
#identical returned 'TRUE', therefore the values are the same
identical( example_descriptiveSTATS, example_descriptiveSTATS2 )
```

### Task #5:
For at least 3 values in a column, please rename so that every value in that column is renamed. For example, suppose I have 20 values of the letter 'e' in one column. Rename those values so that all 20 would show as excellent.

```{r}
#reassigning the values in the 'Colonial' and 'Revolution' collumns to 2 and 3 respectively
#but first, I will display the headers before any changes and order the dataframe such that the entries that will be changes are listed at the top.
head( subQuarrels[ order( -subQuarrels$Colonial ), ] )
head( subQuarrels[ order( -subQuarrels$Revolution ), ] )
#now to change the values
subQuarrels$Colonial[ subQuarrels$Colonial == 1 ] <- 2
subQuarrels$Revolution[ subQuarrels$Revolution == 1 ] <- 3

```

### Task #6:
Display enough rows to see examples of all of steps 1-5 above.

```{r}
#redisplay the header ordered such that we can varify the change
head( subQuarrels[ order( -subQuarrels$Colonial ), ] )
head( subQuarrels[ order( -subQuarrels$Revolution ), ] )
```

### BONUS!
place the original .csv in a github file and have R read from the link. This will be a very useful skill as you progress in your data science education and career.

```{r}
#will start by writing the data
write.csv( Quarrels, file ='Quarrels.csv', row.names = FALSE)
#Great! I see it in my directory, so now i'll beam it up to git.....
#......
#That worked, what a time to be alive.
library(RCurl)
myURL <- 'https://raw.githubusercontent.com/SmilodonCub/MSDS2020_Bridge/master/Quarrels.csv'
gitQuarrels <- read.csv( url( myURL ) )
head( gitQuarrels )
#there is a god. & it's name is StackOverflow
```


